% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%

\documentclass{acm_proc_article-sp}

\begin{document}

\title{The~Framework for~Accessible Applications:
Text-Based Case for~Blind People}
\subtitle{}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{1} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Michael Pozhidaev\\
       \affaddr{National Research Tomsk State University, Computer Science Department}\\
       \affaddr{Lenina Avenue, 36, 634050}\\
       \affaddr{Tomsk, Russia}\\
       \email{msp@altlinux.org}
}
\date{23 October 2014}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}

This~paper offers a~Java framework for~creating accessible applications for~blind and visually impaired people 
as part of~a~proposed general conception
based on~the~maximum use of~objects filled with~text data only.
It offers new types of~applications more easily recognizable by~disabled persons,
helping them to do their work faster and more comfortably.
Strong and weak points are analyzed.
The~published prototype of~the~proposed platform is~described 
as~well as~the~conclusions of~the~performed experiments.
The~prototype is~implemented on~Java~SE and wrapped by~a~GNU/Linux environment 
as~a~bootable ISO-image.

\end{abstract}

\category{D.2.2}{Software Engineering}{Design Tools and Techniques}[User interfaces]
\terms{Experimentation}

\keywords{accessibility,
blind people,
user interfaces,
usability,
Java,
API}

\sloppy

\section{Introduction}

During last decade we have seen the~process of~diversification of~information technologies 
which people use in~their everyday lives.
If during the~mid 2000-s almost all tasks were performed using Microsoft Windows and desktops only,
today's solutions  are very different, regardless of whether we are speaking about software platforms 
(Mac~OS~X, iOs, Android are have a wider distribution and very often can be considered as~real competitors to~Windows)
or about ways of~interaction with PC's or gadgets (multitouch as a~replacement for~the~mouse).
In the meantime, blind and visually impaired people (from here on in ``blind people'')
typically still are~not able to~enjoy all the advantages of~this process,
mostly because the~accessibility technologies are just add-ons 
to~interfaces initially designed for~sighted users.
This kind of~solution can be~considered only a~partial solution 
since it offers the~general ability to~do the~same things as~everybody
but sacrifices the~efficiency and level of~personal comfort.

Each time it~is necessary to~create an~application designed and highly accessible for~blind people 
developers always create a usual application for~graphical user interface (GUI).
Their main distinguishing characteristics aren't more than 
features to~enable the ability to~contrast colours and adjust font sizes.
These functions are only measures for~users with~partial sight.
If somebody has never dealt with~GUI as~a~sighted user   
the~complete and proper understanding is~generally impossible (especially for~senior users).
Respecting the increasing requirements for~an~accessible environment we would  like to~propose 
a~platform designed completely for~blind people 
which could be a~stable solution for~the~problems mentioned above.

Speed, comfort and good understandability  usually mean just one thing; keeping things simple.
The~question is~what kind of~user interface (UI) could be  simple and at~the~same time functionally sufficient.
The~suggested solution to~an~interface for~blind  people
is, on~one hand, simple enough and, on~the~other hand,
could provide  the~replacement for~most GUI~widgets.
We will discuss the pros and cons, 
and discover how visualization and speech output can mutually supplement each other.
The~proposed platform was published on prototypes to~demonstrate 
implementation on~Java. ##This sentence is a bit confusing. Please go into detail or reorder your sentence structure.##
It consists of~the~core maintaining launched applications,
new~interface implementation and a~set of~Java classes making API for~creating new~applications.
##Are you trying to say simply that it contains all of the core applications and interfaces for creating new applications with Java?##
This~environment could be~launched on~any OS with the Java virtual machine 
but we would like to~see it as~a~complete OS on~the~Linux kernel 
(published prototypes are prepared in~this way).

\section{Text-based environment fundamentals}
\subsection{General conception}

First of~all, we would like to~describe an~alternative solution  
to~ensure it~is really suitable and comfortable the~blind.
There is some prior experience in~this area described below and we would like   to~respect it.
##There are several existing applications for the blind and we would like to address them below.##
The~new~environment enable speech output and some visual representation 
without involving any~screen reading software.
The~picture on~the~screen remains very important because it is~used by~users with~low vision, 
though it plays only a very supplementary role now.
We~suggest splitting the~screen into several rectangular areas, 
and filling its~entire space with some tiles. 
Each tile shows some textual data related to~the~corresponding object 
and is closely associated with~speech output.
Font size and colour should be~easily adjustable to~suit a~particular user's requirements.

The~method of~how to~constructing a~comfortable and sufficient speech representation for~all~objects needed for~work
is~one of~the~main goals of~our~research.
It~is~described in~detail below but, roughly speaking,
we~are considering all~of~the~usual GUI widgets  in~a~way where no~graphical and visual data is~involved in~their representation.
Fortunately, there are only a~few cases when it~is~impossible (e.g.~the~web~page cannot be~described in~words without lost of~entirety).
Let's~consider, for~example, a~text~edit box: 
there is~nothing graphical about~it, 
but the~selection of~some part of~the~text with a mouse is~performed through choosing some other colour 
and this happens~by using visual information.
Hence, we must invent something else for~text selection but there is no~need to~create anything new for~text~input itself.

The~environment is~designed for~distribution with~some standard set~of~applications 
prepared and organized in~a~new~way
(it isn't interesting to~have any new~platform for~new~applications
if you still have~to~use usual GUI for~all~other tasks, 
only the~fully consistent environment could be~attractive##Please revise. Difficult to understand exact meaning.##).
At~first glance it~seems that we~will face a tremendous number of~new functions to~implement, 
but actually there is a~wide range of~libraries for~Java, distributed under public licenses;
so there is no need to do anything from~scratch.
We should simply prepare a~new~interface for~them with~our~new~platform.

\subsection{Known experience}

Usually blind people use screen readers launched on~a~GUI, 
but we will not review them as we~are interested only in~software designed as~``audio desktops''.
There are currently three solutions which can be considered as providing a speech environment without screen reading.
All~of~them accomplish this in~different ways.
They are Emacspeak \cite{emacspeak}, 
Dolphin Guide \cite{dolphineguide} and 
Adriane \cite{adriane}. 
Emacspeak certainly had a~major impact on the~outcome of~this~paper.
It~is an~add-on for~the~popular text editor GNU~Emacs \cite{emacs}
which has a~lot of~additional features covering areas significantly wider than just usual text editing, 
such~as file management, mail reading, calendar etc. 
The~main advantage of~Emacspeak is that with some proper training 
it can help users to~be~highly efficient(almost to the level of~a~sighted user).
We~are wanted to~keep all~of~the~positive parts of~that interesting experience.
However, unfortunately, it~has a~lot of~restrictions 
existing mostly as~a~consequence of~its add-on nature 
(GNU~Emacs has nothing close to~applications##Please explain##). 
All~weak points  of~emacspeak have been analyzed  in~our previous publication \cite{luwrain1} in~detail.

Dolphine Guide is a~high-level screen reader with a~lot of~additional information about user interface 
for~a~fixed number of~particular applications.
It obscures interaction with these applications and replaces it with its own 
very user-friendly environment.
Although Dolphin Guide is able to~be a~solution for~inexperienced users 
it cannot be considered as~a~flexible environment providing efficiency for users 
because it handles only a~fixed number of use cases.
In~addition, Dolphin Guide needs a~complete OS (Microsoft Windows) to~work.

These problems are also inherent in~Adriane except 
that it is based on~a~Gnu/Linux distribution.
It involves a~number of~applications for~various tasks 
which were wrapped with speech-enabled interactions.
Adriane cannot be considered as~a~platform for~constructing new accessible applications 
because it hasn't any strict conception##Please explain.## (although its interface could~be considered as~quite consistent)
and the~components it uses are based on~the~Dialog \cite{dialog} utility and Bash scripts \cite{bash}.

\subsection{Pros and Cons}

Let~us consider all the reasons why our application makes sense and look at~arguments for~why  it could have some weak characteristics.
To respond to criticism, which we may expect, 
we~will give additional comments for~why we don't take them seriously and don't think these criticisms are crucial. 
As~a~stronger approach, everything is~written 
like it~would be~as~a~complete OS, and we~will also look at why
the~case of~running in~an~environment on~Microsoft Windows 
is~significantly weaker. 

The~statements supporting  our proposal  could be as~follows: 

\begin{enumerate}

\item {
Users no longer need to~struggle with~GUI and that saves a~lot~of time, increasing efficiency. 
GUI is a~very unsuitable solution 
because in~complicated cases blind users should reference a~screen structure which they don't see.
On~desktops and laptops, GUI is~useful only with~a~mouse,
which is inaccessible for~blind users.
With tablet~computers, where multitouch is~popular, blind users should 
be able to~touch the~screen in~same position multiple times.
That is~easy to~do  having visual information, 
but in~our~speech-enabled case it~turns out as a~very time consuming procedure. 
In~the~meantime, the presence of~GUI is a~completely artificial problem
because it is~created by other people##What do you mean by other people?## and isn't something essential for~interaction with~a~PC.
}

\item {
A~new~type of~interface could be~easier  to~understand 
for~blind users with a~lack of~experience with computers, especially seniors.
With~GUI it~is necessary to~have a~proper understanding of~what it is.
That is not a~problem for~users who have~previously used a~PC as a~sighted user.
But for~those who have never seen a~computer screen such understanding 
becomes a~really serious problem.
}

\item {
The general conception and suggested implementation could be~a~platform 
useful on~mobile laptops as~well as~on embedded devices based on~the~ARM platform \cite{arm}.
In~conjunction with~previous term ##Please explain. Do you want to say the ‘previous sentence’?## this could have social value as~it
would now give access to~a~wide range of~digital services for~disabled people.
This makes it a~solution with social value.
}

\item {
For~experienced users there are certain tasks that~can~be uncovered with existing solutions.
Usually this is~related to~speedy software development,
preparing materials with~Latex \cite{latex} or Lilypond \cite{lilypond}.
A new~approach could be~well adjusted for~these tasks.
}

\end{enumerate}

Criticism of~our~proposal  could be as~follows:

\begin{enumerate}

\item {
It~is in~an~isolated environment.
This makes blind people more isolated from society 
and blocks their access to~software not~included in~the~new~system.
}

\item {
This proposal requires creating some number of~already existing applications,
e.~g. for~mail reading, news~reading etc.
}

\item {
Not all~applications could be~reconsidered with the~proposed conception.
For~example, a~web-browser could be~accessible only with~the~screen reading approach.
}

\end{enumerate}

Our responses to such criticism would be~as~follows:

\begin{enumerate}

\item {About isolation:}
\begin{itemize}
\item {
Our~conception is best suited to~be a~user accompaniment preferably on~mobile computers,
while existing desktop systems remain available 
and everybody is still able to~use a~general purpose OS like Microsoft Windows or Apple Moc~OS~X on~it.
There is no need to have a~specific single universal computer (in~contrast with~mobile phone).
If it~is launched on~Microsoft Windows. Java will be~able to~make this happens.
}
\item {
The~system isn't totally isolated.
It can provide access to~command line utilities 
through Bash or some other shell.
In~fact, command line utilities are one~of~the~most accessible ways for~interaction for~blind users.
Although, of~course, it is suited mostly for~experienced users. 
}
\item {
There is~a~significant gap between potentially possible and really available features for~blind users.
With~GUI we~may think that they~are not~isolated but it doesn't mean that 
we are really able to~do everything that is~needed.
}
\end{itemize}

\item {About creating new software:}
\begin{itemize}
\item {
There is~no~need to~develop anything from~scratch.
We are~working in~the~area of~Free and Open source software,
so we~can use a~lot of~libraries already created for~Java for~our implementation.
}
\item {
According to~the~Pareto rule \cite{pareto},
80\% of~software needs are covered by~approximately 20\% of~the~features in the software.
Therefore, we may expect that there is some appropriate level of~functionality which would be~sufficient for~most tasks.
}
\end{itemize}

\item {About exceptions:
yes, there are some exceptions, but we can include their workarounds into our system as~exceptions.
Speaking about the web-browser, we can take Chromium \cite{chromium} with the~ChromeVox extension \cite{chromevox}
which will run outside of~our environment and that will not bring any inconvenience to~the~user.
Generally, the~complete OS should be some sort of~``hybrid'' system 
if it is based on~the~proposed platform. 
Some popular applications, like photo editing or computer aided design will never be~required 
because they~are pointless without visual information.
}
\end{enumerate}

\section{The framework description and Luwrain project}

In~this section we will consider all~of~the~valuable details  
of~the~Luwrain project which we~proposed as~an~implementation of~the~conception and the~framework described above.
Although Luwrain includes some research and experimental tasks 
we intended  to~get a~stable product and fully functional OS based on~it 
suitable for~developers as~well as~for~a~wide range of~consumers with~sight restrictions.
Anybody who is interested in~getting a~complete understanding of~what Luwrain is and how it~works 
is~welcomed to~try current prototypes freely published on~the~corresponding website at http://luwrain.org/.

We guess that it is necessary to~describe this work from~various points of~view 
as~we~are speaking about not just a~theory but also about an~exact technical approach.
Most things described in~the~sections below should be~reflected  in~the~Luwrain API
as~a~set of~Java classes.
The~Luwrain classes provide various levels of~customization.
For~example, the~class for~list~view needs only a~set of~items 
and does all tasks according to~the~conception. 
However, if~a~developer would like to~prepare his/her own controls, he/she has everything needed to~do~that.
Each application for~Luwrain should be~distributed in~the~form of~.jar file 
as~is usual for~Java libraries.

\subsection{Accessible controls}

The~main requirement imposed on~a~set of~controls 
is to~have functionality equal to~that of~GUI.
Of~course, except obviously, the~requirement of~being~fully accessible for~blind users.
Speaking about a~control'' we mean here items 
such~as text edits, list~views, menus, tables, forms etc.
Forms include various things, like edits, check~boxes, some custom controls and so on.
Every class of~a~control object should generate output for~the~screen and for~speech simultaneously and in~such way 
that output for~speech should be~fully sufficient for~any kind of~work while 
output for~the~screen plays only a~supplementary role.
One additional requirement for~any type of~a~control 
is~providing access to~any part of~the object 
without potentially  inaccessible information.
This is usually very relevant for~exploring the~spelling of~any string  in~a~letter-by-letter manner.
A~lot of~users are unable to~completely rely on~the~pronunciation of~a~speech synthesizer.

\begin{description}

\item[Text edits.]
The~text edits (both single-line and multiline) speak the~letter under the~cursor on~left-right movements 
and speak the~line holding the~cursor on~up-down movements.
On~typing any~letter, the~letter should also be~spoken.
The~important question is~the~selection of~some text fragment.
We~guess that the~most convenient way is~to~set a~special point under the~current cursor position to~mark the~start of~the~region,
then go~to~some other position marking the~end~of~the~region and do one of~the~required operations (copy, cut or delete).
It~is~a~good idea to~pronounce a~text fragment being cut, copied, deleted or~pasted.
On~reaching the~bounds of~a~text area, a~corresponding notification should be~issued.

\item[List views and menus.]  
All~types of~items of~enumeration should  have a~cursor 
marking not~only a~particular line, but also being free to~point to~any~character on~this line.
This is necessary as~mentioned before for~exploring the~spelling of~the~item.
On~up-down movements   a~new~item of~text should be~spoken and~the~cursor should go~to~the~beginning of~the~new line. 
Every line should always have one additional empty line (otherwise, how could we know the~text of~an~item if~it is single and there is no way to~go up-down?).

\item[Tree view.]
Tree views are also possible.
We~can treat them as~some~sort of~extension of~a~list view.
If some particular item has children, it gets a~plus or a~minus sign for~screen representation 
and a~corresponding speech notification is~added.
Pressing the~enter button on~such an~item consistently expands or collapses its subitems.
The~level of~the~item can be~reflected by~the~indentation on~the~screen and corresponding speech suffixes or prefixes.

\item[Forms.]
Forms imply a~set of~various controls such~as text edits, check~boxes, drop~down lists etc.
There are no~problems with them if~the~constructed forms place each control on~a~separate line and adds a~corresponding text prefix 
designating the~name of~the~control.
All~lists should be~drop~down and the~item selection should be~carried~out through additional popup areas (see below).
There is~one noticeable limitation: each form can contain only one multiline edit 
and it always should be~placed at~the~bottom of~the form filling the~entire space below all~other controls.
For~example, such an~approach is~selected for~an~area with the purpose of~composing a~mail message.
It~has a~recipient address on~the~first line, 
and then the~subject on~second, some additional fields,
but below all~of~them there is a~multiline edit for~message text. 
\end{description}

We~described some~the~most important controls for~illustration but, of~course, not~all~of~them.
Others can be~reconsidered in~the~same way as~these. 

There is~one~rather serious problem often faced during work on~accessibility technologies:
there is an~unpredictable amount of~information 
needed by~a~user in~different situations.
When a~user explores a~structure of~a~tree view he/she should get as~much information as possible, 
but if he/she looks for~some particular known item only  the~name is~required to~be~spoken.
We suggest to~use alt keys like Ctrl or Alt on~the~keyboard to switch the~mode of~output.
For~example, holding the~Ctrl key always skips all~supplementary information, saving only the~items names. 

We should mention one additional trick usually considered as~rather popular and useful.
Whereas all controls (lists, texts, trees or forms  can be~represented in~text form in~one or an other way, 
a~feature to~quickly search some~text substring can significantly increase efficiency.
For~forms no~matter  whether this substring appears  in~editable areas or just in control names,
it~should be,
so that users easily understand what it~is by~themselves.

\subsection{Applications and tiling}

Each application in~Luwrain gathers several controls.
Their number and types are defined completely by~the~purpose of~the~application.
For~example, a~mail reader should  consist~of three controls:
tree view with mail groups (Inbox, Sent Items, etc),
the~list of~mail messages inside a~particular group and the~text of~a~particular message.
On~the~screen they should be~placed in~the~same way as~in~a~usual GUI mail client 
but for~users who work through speech  only there shouldn't be any association 
between the~objects location on the~screen and their behaviour.
He/she just has to mind that there are three objects and it has the~proper way to~switch between them.

That could be easily achieved if~the~environment implementation 
takes complete care of~the~calculation of~a~control's position on~the~screen.
We can even think  that there is no need to~have a~way to~choose the~position manually,
it~is enough to~have a~sufficient algorithm giving a~suited position automatically in~most cases.
For~that purpose we would like to~suggest one of~them.
It~would~take a~tree of~tiles, 
each node of~it has references to~the~two children and a~boolean attribute 
whether this node implies dividing  ina~horizontal or vertical way.
Given screen width and height, it calculates the~position of each tile on~the~screen.

\begin{enumerate}

\item {
Performing Depth-first search on~a~tiles tree and~calculating 
how many leaves has each branch of~each node.
}

\item {
As~a~recursive procedure, do the~following steps:
}

\begin{enumerate}

\item{
Call the~procedure providing the~screen width and height as~well as~the~root of~the~tree. 
}

\item {
Indicate if~the provided node is a~leaf assigned to~it received screen position.
}

\item {
Perform the~dividing of~a~received screen area into two~parts 
in~proportion, how many leaves there are under each branch, and handle dividing direction (horizontally or vertically).
After that perform a~call of~the~procedure for~each branch providing the~obtained positions.
}
\end{enumerate}
\end{enumerate}

According to~our experience, this~procedure yields rather good dividing for~each application.
It~is necessary to~describe what an~application is in~Luwrain design.
First of~all, we would like to~note that the~term ``application'' doesn't reflect the~exact nature of~the~implied object.
Very likely it would be~better to~call it an~``applet'' or ``add-on'' because Luwrain applications 
are executed in~the~same process as~the~environment itself (although they~are able to~initiate separate threats##Threats or threads?##) and share with the~environment the~same memory address space.

The~applications are Java classes where objects are registered in~the~corresponding manager.
There could be~multiple instances of~a~particular application (e.~g., user can launch several file managers)
and in~opposite##Why in opposite? Confusing. On the other hand?## there is a~special technique to~prevent some others applications from being launched twice (e.~g., multiple copies of~mail fetching application are pointless).
In~each case, currently only one application can be~shown on~screen 
and it is the~one which is considered to~be~active.
We are thinking about a~special type of~application visible permanently 
(e.~g., for~displaying news feeds)
but we still are~not~sure whether it~is~really necessary or~not.
Switching between applications is~performed easily  and quickly with the~Alt+Tab key combination.

With Luwrain distribution comes a~set of~standard applications.
They are: 
double-sided file manager,
extendable text editor,
mail reader,
news reader,
terminal,
media player,
the~application for~office documents preview,
personal scheduler,
calendar,
address book etc.
Some sort of~applications can be~provided in~the~form of~extensions
and we would like to~see them as~a~part of~community-driven activity.
They are mostly clients for~popular websites, like 
Twitter,
Yandex and Google services,
payment systems, 
social networks etc.

It~is necessary to~especially emphasize the~question of~clients 
for~digital government services.
Their presence in~any~assistive  technologies could have significant social value.
The~question of~an~accessible alternative for~office applications remains very arguable.
Office document exchange is~very active and 
there~is no~doubt that corresponding applications are~needed,
but full functionality with Microsoft Office is needless.
With~Luwrain we~intend to~provide such tools, 
but the~set of~their features is~defined by~corresponding Java libraries (see below).
Office document representation in~text form should be~worked~out, 
but full page rendering (e.~g., for~printing) apparently remains impossible.
As~a~partial measure we suggest to~use non-wysiwyg alternatives, such~as Latex, since they are highly accessible,
 although they require some training and experience. 

\subsection{Events dispatching and popup areas}

The various event dispatching techniques, very likely, are an~essential part of~any UI implementation
and Luwrain isn't an~exception in~this sense.
Luwrain has several types of~events with corresponding rules of~their routing.
They bring information about user actions,
notifications about changes in~the~environment and 
do multithreading synchronization.
The~last is~a~very important feature,
Luwrain allows application developers to initiate as many execution threads as they need, 
but all interaction with~the~Luwrain core should be done in~a~multithreading-safe way
and Luwrain provides some features for~that.

Actually, there is nothing  to~describe in~detail 
except one thing: our environment has special types of~areas which applications are able to~show.
They are the~so called ``popup'' areas.
Their main distinguishing characteristics  is that they can be shown as one method call, 
which ends only on~the~closing of~a~corresponding area.
Since the~environment carries~out in~one single thread, 
this causes some difficulties because the~popup method call (usually placed in~some event handling code)
freezes the~entire event loop execution.
This~problem is~solved by~implementing multiple event loop instances.
The~first of~them is a~main environment event loop and each new loop is~launched for~every new popup area.
We can think of~it~as a~very usual approach in~UI design.

With~popup areas we can~show various dialogs and menus,
continuing the~execution depending on~user choice.
For~example, a~couple of~popup areas have system-wide meaning.
The~first of~them is~the~main menu which is~an~idea very close to~the~``Start'' menu in~Microsoft Windows.
The~other one is~a~command line always accessible with~the~Alt+X key combination.
With this prompt a~user is able to~launch  a~particular application or do some action with system-wide meaning.
This feature is~very useful when it~is necessary to~do something in~a~noisy environment.
For~example, in~some cases it~is easier to~press the~Alt+x and type ``mail'' than to~open the~main menu and listen to~its items.
This~idea of~Alt+X command prompt was adopted from~Gnu~Emacs but with some modifications.

\subsection{Why Java?}

Luwrain is~implemented mostly on~Java.
Its environment is~executed completely inside of~the~Java virtual machine.
The~main reasons why we use Java is the~large variety of~existing Java libraries
and Java is currently a~common language for~any kind of~framework and platform.
Speaking about libraries, for~instance, if  we are creating a~mail client it isn't necessary to~write all the~protocol parsers,
anyone can just use Javamail library \cite{javamail}.
Actually the~number of~involved libraries is~relatively large,
so we use Apache~POI \cite{poi} for~office documents format processing 
(exactly a~functionality of~this library defines how good documents support could be),
Rome \cite{rome} for~RSS parsing,  
and many others.

But the~reasons aren't limited only because of~libraries.
Java has a~rather stable API which is changed very carefully and moderately 
(thoughtless API changing in~our opinion is~one of~the~biggest problems in~the~world of~Open source libraries).
In~addition, \cite{javaspeed} 
the~current Java speed of~execution is~comparable to~the~speed of~C++
and that is~a~rather good result
(sometimes benchmark resources offers information that Java gives 80\% overhead over C++ time).

Some questions related to~Java remain unsolved. 
There~are some legal concerns, we can see that corporations can sue developers over Java \cite{oraclesue}.
Next, it is unclear if we could build something on Dalvik (or on~coming ART) \cite{dalvik}
which looks more efficient for~ARM devices.
Hopefully, these questions could be~solved in~future.

\subsection{System-level services}

If we are speaking about a~complete OS, 
we should keep~in~mind various services for~maintaining network connections and other system tasks, 
as~well as~the~way of~interaction between these services and the~UI inside of~the~Java virtual machine.
Current experience in~the~GNU/Linux world demonstrates a~tendency towards  
D-Bus \cite{dbus} as~a~tool for~interprocess communication (IPC).
Java has a~corresponding interface to~use D-Bus as~well,
therefore, we should just choose projects 
which provide the~necessary functionality with the~D-Bus~interface.
Fortunately, they are present for~almost~all~tasks:

\begin{enumerate} 

\item{
Network manager \cite{nm} for~manipulating network connections
}

\item {
Udisks \cite{udisks} for~removable media management
}

\item {
VoiceMan \cite{voiceman} for~speech output
}

\end{enumerate}

Network Manager and Udisks are well-known projects,
speech server VoiceMan has been developed earlier as~a~part of~the~Luwrain project, 
but it is implemented as~a~system service on~C++.
Currently VoiceMan takes text to~speach through the~inet socket,
but it is just a~temporary measure. 

An~actively discussed idea of~the~Systemd service \cite{systemd}
proposed by~Lenard Poettering 
potentially could be~nicely integrated into~the~system we~are discussing.
Installation on~a~hard drive can be~performed by~a~blind person without any~sighted help 
using the~live system cloning technique \cite{livecdclone}
The~main window of~the~Java environment is~shown with X.org server \cite{xorg}
using a~custom lightweight window manager.

If`Luwrain is~launched on~Microsoft Windows these features will be~inaccessible 
or redirected to~corresponding Windows components.

\section{Obtained experience and testing}##I’m not sure if this is the right title for this paragraph(TESTING AND RESULTS)## 

The~experience we already have consists of~two~parts: 
a general conception of~testing done in~the~Emacspeak environment 
and user feedback collected on~the~publication of~the~first Luwrain prototype.

Using the~proposed conception as~it was~implemented in~Emacspeak was really successful 
(the~author of~this paper successfully graduated university and has~done his Ph.D. thesis using it on a machine).
The~efficiency is~really high and handles some tricky operations,
like an~installation without sighted help.
In~the~meantime, it~is still difficult to~share this~experience with other people 
because using GNU~Emacs and Emacspeak requires a~lot of~technical knowledge.
Some design problems of~GNU~Emacs (e.~g., absence of~an~application)
don't allow us to~consider it~as~a~platform satisfying modern trends for~~popular products.

The~first Luwrain prototype published in~the~form of~a~bootable ISO-image 
was presented on~March 1st, 2014.
Feedback was obtained from~two categories of~users: from newbies and from~experienced users.
The~feedback from~newbies is~more valuable because our~system should
match the~expectations of~wide~range of~users.
With very short prior comments##Do you want to say instructions?## users easily understood what they should do if~they would reach some particular position 
as~well as~to~use some application or open an~object.
In~the~future these instructions can be~offered as~brief guide on~system startup.
Experienced users wanted to~treat the~potential success or failure of~the~system 
as~being~highly dependent on~the~ability to~get a~first stable release,
because this~work obviously requires a~lot~of development resources.

After discussions in~foreign communities, 
the~authors may expect that there could be~some interest 
in~solving accessibility problems specifically through~such specialized environments.

\section{Conclusions}

We~have described all~the~basic questions related 
to~a~special framework for~developing applications for~blind and visually impaired people.
If~this product could be implemented completely,
blind users would get a~free tool for~most of~their everyday operations.
With it they could read and write mail, track news, 
listen to~music and books, etc.
All of these things could~become available easily.
We have omitted the~details of~possible user perception 
because with this paper the~description is~focused on~Luwrain as~a~framework for~creating accessible applications.

The~suggested approach should be~considered very carefully because  accessibility technologies 
always carry some difficulties some amount of~things impossible totally.
This~fact is~quite obvious, 
for~example, when we see that there are~no technologies (and very likely will not appear  for~observable future)
which could be able to~describe in~words any~given picture.
Blind people never will~be able to~do computer aided design and some other tasks in~a~visual nature.
We should look for~sufficient solutions, but not~perfect ones.
With this reality, the~project authors are strongly convinced the~proposed conception could solve some rather tough tasks.
To~obtain it we spent more than ten years in~research including a~large number of~experiments.

It~is~necessary to~properly discuss our arguments against GUI.
A text-based environment should not be~taken as~a~competitor to~GUI.
In~some circumstances the~existing  screen reading solutions for~Microsoft Windows, GNOME or Mac~OS~X
can be the~only possible solution.
For~example, if~somebody must use the~same applications as~everybody in~school or university,
then this would be the case.
Surely, the~solutions in~style of~audio desktops 
should be~aimed at~the~sectors left uncovered with~screen readers.##This sentence is confusing. Please revise.##

We~are still speaking about work which is~just at~the~phase of~representing first prototypes.
And the~number one goal is~to~get it~finished.
The~real value of~this conception can be~measured only by~the~probability of~doing~that.
If~we~are able to~solve some technical questions, 
other things, related to~the~available models for~such work, still remain obscured.
On~the~one~hand, this project is Free and Open source, meaning it is non-profitable,
on~the~other hand, it~could have some social value. 
The~question whether there could be~development models,  more suitable for~this~combination,  or not 
isn't less important than any~technical discussion.##This sentence is confusing, please revise.##

%%\bibliographystyle{abbrv}
%%\bibliography{pozhidaev-luwrain}

\begin{thebibliography}{10}

\bibitem{poi}
D.~V. Ajay~Vohra.
\newblock Use jakarta poi to generate excel spreadsheets from xml documents.
\newblock
  http://www.javaworld.com/article/2076189/enterprise-java/book-excerpt--conve%
rting-xml-to-spreadsheet--and-vice-versa.html, 2006.
\newblock Retrieved May 27 2014.

\bibitem{dolphineguide}
Dolphin Computer Access Ltd.
\newblock Dolphin guide.
\newblock http://www.yourdolphin.com/productdetail.asp?id=30, 2014.
\newblock Retrieved May 27 2014.

\bibitem{rome}
M.~Fortner.
\newblock All roads lead to rome.
\newblock http://www.jroller.com/phidias/entry/ all\_roads\_lead\_to\_rome,
  2009.
\newblock Retrieved May 27 2014.

\bibitem{javamail}
E.~R. Harold.
\newblock {\em JavaMail API}.
\newblock O'Reilly Media, 2013.

\bibitem{dalvik}
D.~Helleberg.
\newblock Android internals: Art in practice.
\newblock
  http://blog.dominik-helleberg.de/2014/01/28/android-internals-art-in-practic%
e/, 2014.
\newblock Retrieved May 27 2014.

\bibitem{oraclesue}
B.~Kendall.
\newblock Oracle wins ruling in case against google over java.
\newblock
  http://online.wsj.com/articles/court-says-oracle-software-code-entitled-to-c%
opyright-protection-1399652818, 2014.
\newblock Retrieved May 27 2014.

\bibitem{lilypond}
P.~Kirn.
\newblock Lilypond: Free, beautiful music notation engraving for anyone.
\newblock
  http://createdigitalmusic.com/2010/05/14/lilypond-free-beautiful-music-notat%
ion-engraving-for-anyone, 2014.
\newblock Retrieved May 27 2014.

\bibitem{adriane}
K.~Knopper.
\newblock Adriane --- audio desktop reference implementation and networking
  environment.
\newblock http://www.knopper.net/knoppix-adriane/index-en.html, 2014.
\newblock Retrieved May 27 2014.

\bibitem{pareto}
R.~Koch.
\newblock {\em The 80/20 Principal}.
\newblock NICHOLAS BREALEY PUBLISHING, LONDON, 1998.

\bibitem{livecdclone}
N.~Koneri.
\newblock Copy your linux install to a different partition or drive.
\newblock
  http://www.linuxjournal.com/content/copy-your-linux-install-different-partit%
ion-or-drive, 2009.
\newblock Retrieved May 27 2014.

\bibitem{latex}
L.~Lamport.
\newblock {\em LaTeX: A Document Preparation System}.
\newblock Addison-Wesley Publishing Company, 2 edition, 1994.

\bibitem{chromium}
R.~Lerner.
\newblock Switching to chrom(ium).
\newblock http://www.linuxjournal.com/content/switching-chromium, 2013.
\newblock Retrieved May 27 2014.

\bibitem{dbus}
R.~Love.
\newblock Get on the d-bus.
\newblock http://www.linuxjournal.com/article/7744, 2005.
\newblock Retrieved May 27 2014.

\bibitem{bash}
C.~Newham.
\newblock {\em Learning the bash Shell}.
\newblock O'Reilly Media, 3 edition, 2005.

\bibitem{javaspeed}
Oracle~corp.
\newblock Java hotspot™ virtual machine performance enhancements.
\newblock
  http://docs.oracle.com/javase/7/docs/technotes/ guides/vm/performance-enhance%
ments-7.html, 2014.
\newblock Retrieved May 27 2014.

\bibitem{systemd}
L.~Poettering.
\newblock systemd, two years later.
\newblock https://archive.fosdem.org/2013/interviews/2013-lennart-poettering/,
  2013.
\newblock Retrieved May 27 2014.

\bibitem{voiceman}
M.~Pozhidaev.
\newblock Speech server voiceman.
\newblock http://marigostra.com/projects/voiceman/, 2013.
\newblock Retrieved May 27 2014.

\bibitem{luwrain1}
M.~Pozhidaev.
\newblock The text-based environment for blind persons: conception and
  operating system design.
\newblock {\em Mezhdunarodnyj nauchno-issledovatelskij zhurnal}, 2:63--66, Feb.
  2013.

\bibitem{chromevox}
T.V~Raman.
\newblock Chromevox: built-in spoken feedback for chrome os.
\newblock
  http://googlecode.blogspot.com/2011/05/chromevox-built-in-spoken-feedback-fo%
r.html, 2011.
\newblock Retrieved May 27 2014.

\bibitem{emacspeak}
T.V~Raman.
\newblock Emacspeak --- the complete audio desktop.
\newblock http://emacspeak.sourceforge.net/, 2014.
\newblock Retrieved May 27 2014.

\bibitem{xorg}
L.~Shiman.
\newblock X.org foundation releases x window system x11r6.7.
\newblock http://lwn.net/Articles/79302/, 2004.
\newblock Retrieved May 27 2014.

\bibitem{emacs}
R.~Stallman.
\newblock Gnu emacs manual.
\newblock http://www.gnu.org/software/emacs/manual/pdf/ emacs.pdf, 2013.
\newblock Retrieved May 27 2014.

\bibitem{dialog}
J.~Tranter.
\newblock Dialog: An introductory tutorial.
\newblock http://www.linuxjournal.com/article/2807, Sept. 1994.
\newblock Retrieved May 27 2014.

\bibitem{arm}
J.~W. Valvano.
\newblock {\em Embedded Systems: Introduction to ARM Cortex-M Microcontrollers
  cover}.
\newblock CreateSpace Independent Publishing Platform, 2012.

\bibitem{nm}
R.~Yuen.
\newblock Introducing networkmanager.
\newblock http://www.redhat.com/magazine/003jan05/features/ networkmanager/,
  2005.
\newblock Retrieved May 27 2014.

\bibitem{udisks}
D.~Zeuthen.
\newblock Simpler, faster, better.
\newblock http://davidz25.blogspot.com/2012/03/simpler-faster-better.html,
  2012.
\newblock Retrieved May 27 2014.

\end{thebibliography}

\balancecolumns
\end{document}
